---
title: "FPCA over trials"
author: "Xinyu Zhang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load relevant libraries and set up.
```{r}
library(fda)
library(funData)
library(MFPCA)
library(tidyverse)
library(landmarkregUtils)
library(emmeans)
library(lme4)
library(scatterplot3d)
library(plotly)
library(htmlwidgets)

Category.colors <- c(NB = "#CD1076", BB = "#FFD39B" )
afcolors <- c(AF = "darkslategray", MN = "orangered")
```

Load data
```{r}
#ex <- 1 # change according to ex number
#raw_curves <- readRDS(file.path(data_dir, str_c("ex1D", ex, "rds", sep = '.'))) %>% ungroup() %>% 
#  mutate(across(c(curveId, Category), ~ factor(.x)))

df <-  read.csv("productionFullTable.csv")

#change variables into correct data types
df$BB <- as.factor(df$BB)
df$Label <- as.factor(df$Label)
df$block <- as.factor(df$block)
df$gender <- as.factor(df$gender)
df$AF <- as.factor(df$AF)
df$subj <- as.character(df$subj)
df$seq <- as.numeric(df$seq)
df$F0Hz <- as.numeric(df$F0Hz)
df$F1Hz <- as.numeric(df$F1Hz)
df$F2Hz <- as.numeric(df$F2Hz)
df$F2Bark <- as.numeric(df$F2Bark)
df$F2Bark <- as.numeric(df$F2Bark)


#mutate a new sequence number that does not return to zero at the start of every block
df <- df %>% 
  mutate(
    seq = case_when(
      block == "2" ~ seq +54,
      block == "4" ~ seq +107,
      TRUE ~ seq
  )
) %>% 
  rename("biteblock" = "BB", "auditoryfeedback" = "AF")


#df %>%
  #filter(Label == "I") %>%
  #count(subj, seq, Label) %>%
 # filter(n > 0)

df <- df %>%
  filter(block =="2" | block == "1" ) %>%
  #filter(block == "2") %>% 
  filter (Label == "I") %>% #all blocks
  distinct(subj, seq, .keep_all = TRUE)

# nSamples <- 54
# grid <- seq(1, 54, length.out = nSamples)
# 
# adf <- df %>%
#   group_by(subj) %>%
#   arrange(seq) %>%
#   reframe({
#     # extract vectors explicitly
#     x <- seq
#     y <- F2Hz
# 
#     # remove NA pairs
#     keep <- is.finite(x) & is.finite(y)
#     x <- x[keep]
#     y <- y[keep]
# 
#     # remove duplicates in x
#     d <- !duplicated(x)
#     x <- x[d]
#     y <- y[d]
# 
#     # safety check
#     if(length(x) < 2) {
#       return(tibble(seq = grid, F2Hz = NA_real_))
#     }
# 
#     tibble(
#       seq = grid,
#       F2Hz = approx(x = x, y = y, xout = grid, rule = 2)$y
#     )
#   }) %>%
#   ungroup()
```
Plot a few curves
```{r}
# plot a few curves

subset_fileName <- df %>%
  ungroup() %>% 
  distinct(subj) %>%
  slice_sample(n = 20)

curvedf<- df %>% inner_join(subset_fileName, by = "subj")


ggplot(data = df) + #or use df to plot all curves
  aes(x = seq, y = F2Hz, group = subj, color = biteblock, linetype = auditoryfeedback) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values=Category.colors) +
  scale_x_continuous(breaks = c(1, 54, 107, 161))+
  labs(x = "trial number")+
  #mytheme  +
  #theme(legend.position = "bottom")
  theme_light()

```



Compute FPCA
```{r}
# build a funData object (limiting it to only one of the vowels for simplicity)

funF2 <- df %>% 
  long2irregFunData(id = "subj", time = "seq", value = "F2Hz") %>% 
    as.funData()

# Compute FPCA
fpcaF2 <- PACE(funF2)

#df %>% filter(subj == "1")

#proportion of explained variance
fpcaF2$values / sum(fpcaF2$values)

```


Plot PC curves
```{r}
# scores st. dev.
sdFun <- fpcaF2$values %>% sqrt()
# PC curves to be plotted
nPC <- 3
PCcurves <- expand_grid(PC = 1:nPC,
                        fractionOfStDev = seq(-1, 1, by=.25)) %>%
  group_by(PC, fractionOfStDev) %>%
  reframe(
    funData2long1(fpcaF2$mu + fractionOfStDev * sdFun[PC] * fpcaF2$functions[PC],
                  time = "seq", value = "F2Hz")
  )
# Plot
ggplot(PCcurves) +
  aes(x = seq, y = F2Hz, group = fractionOfStDev, color = fractionOfStDev) +
  geom_line() +
  scale_color_gradient2(low = "deeppink2", mid = "grey", high = "burlywood1",
                        breaks = c(-1, 0 , 1)) +
  facet_wrap(~ PC, nrow = 1,
             labeller = labeller(PC = ~ str_glue("PC{.x}"))) +
  labs(color = expression(frac(s[k], sigma[k]))) +
  xlab("trials") +
  geom_line(data = PCcurves %>% filter(fractionOfStDev == 0), color = 'black', linewidth = 1.2, linetype = "dashed") +
  scale_x_continuous(breaks = c(54, 107, 161))+
  #mytheme +
  theme_minimal()+
  theme(legend.position = "bottom")

```

Plot the mean curve
```{r}
# quick way (not ggplot)
fpcaF2$mu %>% plot()
# nicer way
fpcaF2$mu %>% funData2long1(time = "seq", value = "F2") %>% ggplot() +
  aes(x = seq, y = F2) +
  scale_x_continuous(breaks = c(54, 107, 161))+
  geom_line()
```

Plot PC1
```{r}
fpcaF2$functions[1] %>% funData2long1(time = "time", value = "y") %>% ggplot() +
  aes(x = time, y = y) +
  scale_x_continuous(breaks = c(54, 107, 161))+
  geom_line()
```

Print the PC scores of the first few curves
```{r}
fpcaF2$scores %>% head()
```

Reconstruct the first curve (first row of scores matrix) up to the first two PC scores (first two columns of scores matrix). Compare to original curve.

```{r}

orig_curve <- funF2[2] %>% funData2long1(value = "F2Hz")
rec_curve <- (fpcaF2$mu + fpcaF2$scores[2,1] * fpcaF2$functions[1]+
                fpcaF2$scores[2,2] * fpcaF2$functions[2] 
                ) %>% funData2long1(value = "F2Hz")
# join the curves into one long format tibble and plot
bind_rows(orig_curve %>% mutate(curve = "original"),
          rec_curve %>% mutate(curve = "reconstructed")) %>% 
  ggplot(aes(time, F2Hz, color = curve)) +
  scale_x_continuous(breaks = c(54, 107, 161))+
  geom_line()
  
```

Plot PC scores scatterplot
```{r}
PCscores <- fpcaF2$scores %>%
  `colnames<-`( paste0("s", 1:fpcaF2$npc)) %>%
  # above line means: colnames(fpcaF2$scores) <- paste0("s", 1:fpcaF2$npc) %>% 
  as_tibble() %>%
  bind_cols(df %>% distinct(subj, biteblock, gender, auditoryfeedback), .)


# scatterplot PC scores s1 and s2 by BB
ggplot(PCscores) +
  aes(x = s1, y = s2, color = biteblock) +
  geom_point() +
  scale_color_manual(values=Category.colors) +
  #mytheme +
  #theme(legend.position = "right")
  theme_minimal()


# scatterplot PC scores s1 and s2 by AF
ggplot(PCscores) +
  aes(x = s1, y = s2, shape = biteblock, color = auditoryfeedback) +
  geom_point() +
  scale_color_manual(values= afcolors) +
  #mytheme +
  #theme(legend.position = "right")
  theme_minimal()

af <- Category.colors[as.numeric(PCscores$auditoryfeedback)]
bb <- c(16, 17)
bb <- bb[as.numeric(PCscores$biteblock)]
scatterplot3d(PCscores$s1, PCscores$s2, PCscores$s3, pch = bb, color = af, angle = 55)



s3plot <- plot_ly(
  PCscores, x = ~s1, y = ~s2, z = ~s3, 
  color = ~auditoryfeedback, colors = afcolors
  #color = ~biteblock, colors = Category.colors
  ) %>%
  add_markers() %>%
  layout(
    scene = list(xaxis = list(title = 'PC1 score'),
        yaxis = list(title = 'PC2 score'),
        zaxis = list(title = 'PC3 score'))
        )

saveWidget(s3plot, file = "3dplotAF.html")
```

Pick two random curves from the dataset, call them  `curve1` and `curve2`. Plot the curves. Then mark the corresponding points on the scatterplot.
```{r}
id1 <- 4
id2 <- 35
curve1 <- df %>% filter(subj==id1)
curve2 <- df %>% filter(subj==id2)
ggplot(mapping = aes(seq, F2Hz)) +
  geom_line(data = curve1, color = "magenta") +
  geom_line(data = curve2, color = "chartreuse") 

ggplot(PCscores) +
  aes(x = s1, y = s2) +
  geom_point() +
  geom_point(data = PCscores %>% filter(subj==id1), color = "magenta", size = 3, shape = 18) +
  geom_point(data = PCscores %>% filter(subj==id2), color = "chartreuse", size = 3, shape = 18) +
  #mytheme +
  theme(legend.position = "right")
```

Construct a new curve based on  `curve1` and `curve2`, call it `curveX`. The new curve is obtained by averaging the PC scores of  `curve1` and `curve2`. Consider the first two PC scores. Plot `curveX`'s PC scores on the scatterplot. Plot `curveX` together with  `curve1` and `curve2`.
```{r}
scoresX <- PCscores %>% filter(subj %in% c(id1, id2)) %>% summarise(across(c(s1, s2), mean))
ggplot(PCscores) +
  aes(x = s1, y = s2) +
  geom_point() +
  geom_point(data = PCscores %>% filter(subj==id1), color = "magenta", size = 3, shape = 18) +
  geom_point(data = PCscores %>% filter(subj==id2), color = "chartreuse", size = 3, shape = 18) +
  geom_point(data = scoresX, color = "blue", size = 3, shape = 18) +
  geom_segment(x = PCscores %>% filter(subj==id1) %>% pull(s1),
               xend = PCscores %>% filter(subj==id2) %>% pull(s1),
               y = PCscores %>% filter(subj==id1) %>% pull(s2),
               yend = PCscores %>% filter(subj==id2) %>% pull(s2),
               linetype = "dashed") +
  #mytheme +
  theme(legend.position = "right")

curveX <- (
  fpcaF2$mu + (scoresX %>% pull(s1)) * fpcaF2$functions[1] +
    (scoresX %>% pull(s2)) * fpcaF2$functions[2]
) %>% funData2long1(time = "seq", value = "F2Hz")
ggplot(mapping = aes(seq, F2Hz)) +
  geom_line(data = curve1, color = "magenta") +
  geom_line(data = curve2, color = "chartreuse") +
  geom_line(data = curveX, color = "blue")+
  scale_x_continuous(breaks = c(1,54, 107, 161))
```

Reconstruct the first curve (first row of scores matrix) up to the first two PC scores (first two columns of scores matrix). Compare to original curve.
```{r}
orig_curve <- funF2[1] %>% funData2long1(time = "seq",value = "F2Hz")
rec_curve <- (fpcaF2$mu + fpcaF2$scores[1,1] * fpcaF2$functions[1] + 
                fpcaF2$scores[1,2] * fpcaF2$functions[2]) %>% funData2long1(time = "seq", value = "F2Hz")
# join the curves into one long format tibble and plot
bind_rows(orig_curve %>% mutate(curve = "original"),
          rec_curve %>% mutate(curve = "reconstructed")) %>% 
  ggplot(aes(seq, F2Hz, color = curve)) +
  scale_x_continuous(breaks = c(54, 107, 161))+
  geom_line()
```


Fit a regression model on s1
```{r}
#model for s1
s1mod <- lm (s1 ~ biteblock, data = PCscores)
emmeans(s1mod, pairwise ~biteblock)

#reconstruct predicted curves
emm <- emmeans (s1mod, pairwise ~ biteblock) $emmeans %>% 
  as_tibble() %>% 
  #select(biteblock, emmean) %>% 
  rename(s1 = emmean)

predCurvess1 <- emm %>% 
  group_by(biteblock) %>% 
  reframe(funData2long1(fpcaF2$mu + s1 * fpcaF2$functions[1])) %>% 
  rename(seq = time, F2 = value)

ggplot(predCurvess1) + 
  aes(seq, F2, color = biteblock)+
  geom_line()+
  labs(x = "trial number")+
  scale_color_manual(values = Category.colors) + 
  scale_x_continuous(breaks = c(54, 107, 161))+
  #mytheme +
  theme_light()
```

linear model on PC scores
```{r}
s1lm <- lm(s1 ~  biteblock * auditoryfeedback + gender, data = PCscores)
summary(s1lm)
```


Plot the mean curve, then add one SD of s1 to it, on top of it, substract an SD of S1, to see what the first PC does
```{r}
#the mean curve
meancurve<- fpcaF2$mu %>% funData2long1(time = "seq", value = "F2") #%>% ggplot() +
  #aes(x = seq, y = F2) +
  #scale_x_continuous(breaks = c(54, 107, 161))+
  #geom_line()

#the mean curve with 1sd of s1 added
#plus1sdcurve <- fpcaF2$mu %>%funData2long1(time = "seq", value = "F2") %>% 
  #mutate(F2 = F2 + sd(PCscores$s1))#%>% ggplot() +
  #aes(x = seq, y = F2) +
  #scale_x_continuous(breaks = c(54, 107, 161))+
  #geom_line()


sd1 <- fpcaF2$mu + 1*sd(PCscores$s1)*fpcaF2$functions[1]

plus1sdcurve <- sd1%>% funData2long1(time = "seq", value = "F2") 

sd_1 <- fpcaF2$mu - 1*sd(PCscores$s1)*fpcaF2$functions[1]

minus1sdcurve <- sd_1%>% funData2long1(time = "seq", value = "F2") 

# join the curves into one long format tibble and plot
bind_rows(meancurve %>% mutate(curve = "mean"),
          plus1sdcurve %>% mutate(curve = "+1sd"),
          minus1sdcurve %>% mutate(curve = "-1sd")) %>% 
  ggplot(aes(seq, F2, color = curve)) +
  scale_x_continuous(breaks = c(54, 107, 161))+
  geom_line()

#maybe it shouldn't be the mean curve? Check Gubian
```



