---
title: "Production Analysis"
author: "Xinyu Zhang"
date: "2023-03-27"
output: html_document
---

### read in the list of file names and add a header
```{r}
rcds<- read.csv("fileList.csv", header = FALSE)
colnames(rcds) <- c("filename")
head(rcds)
```

### splitting the different segments in the file name into columns because they provide info about the recording

```{r}
prod <- read.table(text = as.character(rcds$filename), sep = "_")
head(prod)
```
### removing the ".wav" from the final column
```{r}
library(stringr)
library(dplyr)
prod <- prod %>% 
  mutate(V5 =
    str_sub(prod$V5, end=-5)
  )
head(prod)
```

### renaming the column names accordingly
```{r}
newcols <- prod %>% 
  select(-V1)
colnames(newcols)

 
nc<- newcols %>% 
  rename("subj" = "V2",
         "word" = "V3",
         "block" = "V4",
         "seq" = "V5")

head(nc)
```

the naming convention was wrong in the posttest recordings

Such that the order was not condition_sequenceNr but sequenceNr_condition. Fixing this

### separating the posttest recordings and non-posttest recordings
```{r}
library(dplyr)
p <- nc %>%
  filter(seq == "Post")
nrow(p)

np <- nc %>% 
  filter(seq != "Post")
nrow(np)
  
```

### reversing the content of the last two columns in the posttest subset

```{r}
post <- p %>% rename(seq = block,
             block = seq)
head(post)

wposttest <- post %>% relocate(seq, .after = block)
head(wposttest)
```

### row-binding the posttest subset and the non-posttest subset
```{r}
all <- rbind(np, wposttest)
nrow(all)
```

### remove all the practice tokens
```{r}
alldata <- all %>% 
  filter(seq != "practice") %>% 
  filter(block != "practice")
```

```{r}
alldata %>% 
  group_by(subj) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
```



### sequence number should be numeric
```{r}
alldata$seq <- as.numeric(alldata$seq)
```

### previewing the current structure
```{r}
head(alldata)
write.csv(alldata,'recordingInfo.csv', row.names=FALSE)
```

